---
title: "R Notebook"
output:
  html_document:
    df_print: paged
Author: Seth Harrison
---
#Seth Harrison{-}

Download the file MarketShare.csv then write an R Notebook document that will contain the R codes and answers to the following problems. 

```{r}
dat <- read.csv("MarketShare.csv")
```

1. Fit a linear regression model that can be used to predict CustomerSatisfaction based on MarketShare.  Interpret the estimate of the slope and intercept of the model.

```{r}
modCS <- lm(CustomerSatisfaction ~ MarketShare, dat)
summary(modCS)
```

* The model has an intercept of 8.39331, meaning when the value of market share is 0, customer satisfaction is at a level of 8.39331. The  slope of the model is -0.17291, meaning that for each increase in the value of the market share by 1, the customer satisfaction will decrease by -0.17291.

2. Is MarketShare a statistically significant predictor of CustomerSatisfaction?  Justify by citing an appropriate statistic from the regression summary.

* The value of the coefficient of determination, R-Squared, is .8079 meaning that 80.79% of our customer satisfaction variance can be explained using this model. This is a statistically significant model and would be an appropriate model to predict customer satisfaction based on market share.

3. Plot the residuals after fitting the model in #1.  Do the residuals appear to be white noise?  Verify with an ACF plot.

```{r}
plot.ts(modCS$residuals)
acf(modCS$residuals)
```

* Based on the plot of the residuals and the ACF plot, we are led to believe that the residuals follow a white noise distribution.

4. What can we conclude if the residual is a white noise?
* We can conlude that the model fits the data well.



5. Fit a linear regression model that can be used to predict CustomerSatisfaction at time t based on MarketShare at time t-1.  Which is a better predictor of CustomerSatisfaction at time t, is it MarketShare at time t or the lag 1 Marketshare?  Explain.

```{r}
modCSLag <- lm(dat$CustomerSatisfaction[2:15] ~ dat$MarketShare[1:14])
summary(modCSLag)
summary(modCS)
```
* The model predicting customer satisfaction using lag 1 market share is better because it has a R-Squared value of .8677 and an adjusted R-Squared value of .8567, while the model without lag has an R-Squared value of .8079 and adjusted R-Squared value of .7931.

6. Do Problem 2.1 from the textbook.

2.1 A Structural Model For the Johnson & Johnson data, say yt
, shown in Figure 1.1,
let $x_t = log(y_t)$. In this problem, we are going to fit a special type of structural model,
$x_t = T_t + S_t + N_t$ where $T_t$
is a trend component, $S_t$
is a seasonal component, and $N_t$
is noise. In our case, time t is in quarters (1960.00, 1960.25, . . . ) so one unit of time
is a year.

(a) Fit the regression model
\begin{equation}
x_t = \underbrace{β_t}_\text{trend} +\underbrace{α_1 Q_1(t) + α_2 Q_2(t) + α_3 Q_3(t) + α_4 Q_4(t)}_\text{seasonal} + \underbrace{wt|{z}noise}_\text{noise}
\end{equation}
where $Q_i(t) = 1$ if time t corresponds to quarter i = 1, 2, 3, 4, and zero otherwise.
The $Q_i(t)$’s are called indicator variables. We will assume for now that wt
is a Gaussian white noise sequence. Hint: Detailed code is given in Code R.4, the last
example of Section R.4.

```{r}
library(astsa) # SEE THE FOOTNOTE
plot(jj, type="o", ylab="Quarterly Earnings per Share")
trend = time(jj) - 1970 # helps to 'center' time
Q = factor(cycle(jj) ) # make (Q)uarter factors
reg = lm(log(jj)~0 + trend + Q, na.action=NULL) # no intercept
head(model.matrix(reg)) # view the model design matrix
summary(reg)
```


(b) If the model is correct, what is the estimated average annual increase in the logged
earnings per share?

```{r}
reg$coefficients
```

* The average annual increase in earnings per share is .1671722.

(c) If the model is correct, does the average logged earnings rate increase or decrease
from the third quarter to the fourth quarter? And, by what percentage does it
increase or decrease?

```{r}
abs(reg$coefficients[4] -reg$coefficients[5])/ sum(reg$coefficients[4])
```

* The average logged earnings rate decreases by 23.34% between Q3 and Q4.

(d) What happens if you include an intercept term in the model in (a)? Explain why
there was a problem.

```{r}
trend = time(jj) - 1970 # helps to 'center' time
Q = factor(cycle(jj) ) # make (Q)uarter factors
regWithInt = lm(log(jj) ~ trend + Q, na.action=NULL) # no intercept
head(model.matrix(regWithInt)) # view the model design matrix
summary(regWithInt)
```

* When we include an intercept in the data, we no longer have Q1 as a predictor and we have lower R-Squared values.

(e) Graph the data, $x_t$, and superimpose the fitted values, say $x^t$, on the graph. Examine the residuals, $x_t − x^{t}$, and state your conclusions. Does it appear that the model fits the data well (do the residuals look white)?

```{r}
par(mfrow=c(1,2))
plot(log(jj), main="data and fitted value")
lines(fitted(reg), col="blue")
plot(log(jj)-fitted(reg), main="residuals")
```

* It would appear that the residuals are approximately white and that the model fits the data well.(wrong)


